---
title: "Getting Started with rminka"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with rminka}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r carga, include=FALSE}
library(dplyr)
library(jsonlite)
library(httr)
#library(leaflet)
#library(magick)
library(rminka)
library(tibble)
library(knitr)
```

## About

`rminka` is a wrapper for Minka APIs for accessing the observations.


## Installation

The easiest way to get `rminka` is to install from github repository with `pak`: 

```{r install, eval = FALSE, echo= TRUE}
pak::pak("Raiservi/rminka")
```

## Quickstart guide

These functions are designed to be used together. For queries that span multiple years, you can easily loop through the years of interest, run the appropriate function, and then combine the resulting tibbles with `dplyr::bind_rows()`.
> It is recommended to set the `quality` argument to `"research"` in order to 
get more reliable data that has been validated by several contributors.

### Project Queries

A set of complementary  functions to find projects and their associated observations.

**- mnk_proj_byname ** 

   `mnk_proj_byname()`Finds a project's ID using an approximate project name.


```{r project_name, include = TRUE , echo=TRUE}

prj_names <- mnk_proj_byname("2025")

prj_names[,c(1:5)]

```

**- mnk_proj_info **

   `mnk_proj_info()`: Retrieves detailed project information using its known ID.
  
  
```{r project_info, include = TRUE , echo=TRUE}

prj_info <- mnk_proj_info(420)

kable(as.data.frame(prj_info))

```
It´s posible to known the users subscrived in this projects chnging de parameter 'user = TRUE'  and the function gives a list of the id users.

```{r project_info_users, include = TRUE , echo=TRUE}

prj_info <- mnk_proj_info(420, users = TRUE)

prj_info

```
  
**- mnk_proj_obs **
   `mnk_proj_obs()`: Fetches all observations for a specific year within that project.

### User Queries

Functions to find users and their contributed observations.

**- mnk_user_byname ** 
    `mnk_user_byname()`: Finds a user's ID from their approximate login name.
    
```{r user_name, include = TRUE , echo=TRUE}

name <- mnk_user_byname("ramon")

kable(name[,1:4])
```


**- mnk_user_proj **
    `mnk_user_proj()`: Find the projects to which a user has explicitly subscribed    based on their user_ID. You need to known previously the user_id using teh function `mnk_user_byname` 
    
```{r user_prj, include = TRUE , echo=TRUE}
user <- mnk_user_proj(6)

(user[,1:2])
```

    
**- mnk_user_obs **
    `mnk_user_obs()`: Retrieves all observations contributed by that user for a given year ( all the year or only a specific month) from their id_user.

### Place Queries

Functions to find places and retrieve their spatial data.

**- mnk_places_byname ** 
   `mnk_places_byname()`: Finds the ID for a location using an approximate place name.
   
**- mnk_places_sf **  
   `mnk_place_sf()`: Returns the `sf` geometry for a place, ready for plotting with packages like `ggplot2` or `leaflet`.

**d) Observation Queries:** A variety of functions to fetch observation data based on different parameters.

*   `mnk_obs_id()`: Retrieves a single observation's complete data using its unique ID.
*   `mnk_obs()`: Fetches observations based on various parameters for a full year, a specific month, or a single day.
*   `mnk_obs_bydays()`: Retrieves all observations within a date range in the same year.

**e) Auxiliary functions:**  a set of functions with different utilities that complement Minka's observational data and help in processing them when used in other R packages (`vegan`, `dismo`, `labdsv` or others).

*   `get_wrm_tax()`: Retrieves the complete taxonomy and additional data (e.g., marine or terrestrial ) from  [WoRMS](https://www.marinespecies.org/) for a given scientific name.
*   `shrt_name()`: Returns the CEP name (abbreviated scientific name) with separation point, given a scientific name.





`mnk_user_proj` This function retrieves observations from Minka. The text or taxon search can be refined by observation date, record quality and location.



#### Get user id

You can search for observations by either common or scientific name. It will search the entire  database, so the search below will return entries that _mention_ seabream fishes , not just seabream observations.




> Note that `get_minka_obs()` will return 500 observations by default. This can be controlled with the `maxresults` argument.

```{r searchrrr, include = TRUE , echo=TRUE}

# sard_max_results <- get_minka_obs(projectyear = 2021, maxresults = 10)
# 
# sard_max_results[,1:6]
```


#### Taxon search

To return only records of a specific species or taxonomic group, use the `taxon_name` argument. For example, to return observations of anything from the family Sparidae, and restricting the search to the year 2022: 

```{r search2, include = TRUE , echo=TRUE}

# sparidae <- get_minka_obs(taxon_name  = "Sparidae", year = 2022)

# how many unique taxa?

# length(unique(sparidae$scientific_name))

#The species are:

# unique(sparidae$scientific_name)
```

 And to return only the seabream fish observations that also mention the term "banc":


```{r search3, include = TRUE , echo=TRUE}

# sard_banc <- get_minka_obs(taxon_name = "Diplodus", query = "banc")
# 
# count_sard <- nrow(sard_banc)

#count_sard

#the observarions are

# sard_banc[1:count_sard,1:2]

```


#### Bounding box search

You can also search within a bounding box by giving a simple set of coordinates.
The set of coordinates are defined by c( min Long, min Lat, max Long, max Lat)
 
```{r search4, include = TRUE , echo=TRUE, fig.cap = "Map of Salpa salpa observations"}
# 
# bounds_forum <- c(41.407281, 2.224977, 41.409892, 2.228652 )
# 
# diplodus_forum <- get_minka_obs(query = "Diplodus sargus", bounds = bounds_forum, year = 2024, month = 5, maxresults = 7)
# 
# diplodus_forum[1,1:6]
```   
and the representation using leaflet is:

```{r leaflet, include = TRUE, echo=TRUE, fig.pos='center', fig.cap='Leaflet map with the observations'}


#  leaflet() %>%
#          addTiles() %>%
#          setView(lng = 2.226725, lat = 41.408492, zoom = 16.35) %>%
#           addRectangles(lng1 = 2.224977, lat1 =  41.407281,
#           lng2 = 2.228652, lat2 = 41.409892) %>%
#           addMarkers (
#             lng = diplodus_forum$longitude, 
#             lat = diplodus_forum$latitude,
#             label = paste('Id =', as.character(diplodus_forum$id))
#              )
```  

### Other functions

#### Get information and observations by project

You can get all the observations for a project if you know its ID or name as an Minka slug.


```{r search5, include = TRUE , echo=TRUE}

# ## Just get info about a project
# urbamar <- get_minka_obs_project("urbamarbio", type = "info", raw = FALSE)
# urbamar$project_observations_count
```


```{r search6, include = TRUE , echo=TRUE, message=TRUE}
## Now get all the observations for that project
# urbamar_obs <- get_minka_obs_project("biodiversitat-marina-de-les-platges-de-sant-adria-del-besos", type = "observations")

#Geting the first 5 observations

# urbamar_obs[1:5,8:1]
```


#### Get observation details

Detailed information about a specific observation can be retrieved by observation ID. The easiest way to get the ID is from a previous search for that taxon.


```{r search7, include = TRUE , echo=TRUE}
# ident <- sard_banc$id[1]
# 
# ident
# 
# get_minka_obs_id(ident)[1:5]
```


#### Get all observations by user

If you just want all the observations by a user you can download all their observations by user ID. A word of warning though, this can be quite large (easily into the 1000's).


```{r picture, include = TRUE , echo=TRUE, fig.cap='Picture of an user observation',fig.pos='center',fig.height=230}

# user_obs <- get_minka_obs_user('ramonservitje')
# 
# (user_obs)[1:5,1:3]
# 
# #it is possible to view the images with the links in Minka
# 
# url <- user_obs$image_url[412]
# 
# # Read the image
# image <- image_read(url)
# 
# # Show the image
# print(image)
```

