---
title: "Introduction to rminka"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to rminka}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup}
library(rminka)
```
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r carga, include=FALSE}
library(dplyr)
library(jsonlite)
library(httr)
library(leaflet)
#library(magick)
library(rminka)
library(tibble)
library(knitr)
```


This document introduces you to rminka basic set of tools, and shows you how to 
apply them to obtain your desired information.
Once you’ve installed, read vignette("rminka") to learn more.

We will develop an example based on a project: Biomarato Tarragona. 
The example will focus on this project and one of its participants, Ramon Servitje. 
As the example progresses, we will use the different types of functions from the package.


### - Project Queries 

- **<u>mnk_proj_byname</u>**

Initially, only the project name is known, so a search will be performed to find 
the project ID (project_id). The `mnk_proj_byname` function will be used for this purpose. 
A generic search will be conducted using the text "biomarato 2025".



```{r project1, include = TRUE , echo=TRUE}
prj_names <- mnk_proj_byname("biomarato 2025")

prj_names 

# For this example 

prj_names[,c(1:2)]

```
- **<u>mnk_proj_info</u>**
 
   Retrieves detailed project information using its known ID.
```{r project_info, include = TRUE , echo=TRUE}

prj_info <- mnk_proj_info(419)

prj_info

```

It´s posible to known the users subscribed in this projects changing de parameter 
'user = TRUE'  and the function gives a list of the id users.

```{r project_info_proj, include = TRUE , echo=TRUE}

prj_info_us <- mnk_proj_info(419, users = TRUE)

prj_info_us

```

  - **<u>mnk_proj_obs</u>**

Retrieves 

```{r project_obs:proj, include = TRUE , echo=TRUE}

prj_obs <- mnk_proj_obs(419, year= 2025, month=5)

prj_obs[2:14]

```


### - User Queries  
  
 - **<u>mnk_user_byname</u>**

```{r user_name, include = TRUE , echo=TRUE}

user_name <- mnk_user_byname("ramon")

user_name

```

 - **<u>mnk_user_proj</u>**
 
```{r user_project, include = TRUE , echo=TRUE}

user_project <- mnk_user_proj(6)

user_project

```

 - **<u> mnk_user_obs</u>**

```{r user_obs, include = TRUE , echo=TRUE}

user_obs <- mnk_user_obs(user_id=6, year = 2025, month = 8)

user_obs

```


### - Place Queries

 - **<u>mnk_places_byname</u>**

```{r places_byname, include = TRUE , echo=TRUE}

places <- mnk_places_byname("Forum")
places[,1:6]

```

 - **<u>mnk_places_sf</u>**

```{r places_sf, include = TRUE , echo=TRUE}

library(sf)
library(leaflet)


# 1. Downloading the geometry 
place <- mnk_place_sf(253)

# 2. Adding information to the layer  from previous function
place <- place |>
  mutate(
    name = places[2,3],
  )

# 3. Drawing the map
 forum_sf <-leaflet(place) |>
                addProviderTiles("OpenStreetMap", group = "OSM") |>
                addProviderTiles("Esri.WorldImagery", group = "Satélite") |>
                addPolygons(
                  color = "#2c4fb8",
                  weight = 2,
                  opacity = 1,
                  fillOpacity = 0.4,
                  label = ~name, # information added from previous function
                  highlightOptions = highlightOptions(weight = 3, bringToFront = TRUE)
                ) |>
                addLayersControl(baseGroups = c("Satélite", "OSM")) 
 forum_sf

```

 - **<u>mnk_places_obs</u>**
 
```{r places_obs, include = TRUE , echo=TRUE}

library(sf)
library(leaflet)

obs_place <- mnk_place_obs(place_id= 253, year = 2025, month = 2)

obs_place


obs_sf <- obs_place |>
  # 1. Selecting 
  transmute(
    id,
    taxon_name,                    # si no existe, créala antes con mutate(taxon_name = "...")
    observed_on = as.Date(observed_on),
    latitude,
    longitude,
    positional_accuracy
  ) |>
  # 2. elimina filas sin coordenadas
  filter(!is.na(latitude), !is.na(longitude)) |>
  # 3. convierte a puntos WGS84. remove = FALSE mantiene lat/lon como columnas
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326, remove = FALSE)

forum_sf |>
  addMarkers(
    data = obs_sf,
    popup = ~paste0(
      "<b>", taxon_name, "</b><br>",
      "ID: ", id, "<br>",
      "Fecha: ", observed_on
    ),
    clusterOptions = markerClusterOptions()
  ) |>
  addLayersControl(baseGroups = c("Satélite", "OSM"))

```


### - Observation Queries

 - **<u>mnk_obs_id</u>**

```{r obs_id, include = TRUE , echo=TRUE}

obs_id <- mnk_obs_id(id = 553028)

obs_id

```

 - **<u>mnk_obs</u>**
 
```{r obs, include = TRUE , echo=TRUE}

#In this example don´t show messages in console (quiet= TRUE)

obs <- mnk_obs(taxon_name= "Diplodus sargus", year=2025, user_id=6,quiet = TRUE)

obs[,c(1,21,2,10,11,16,27)]

obs

obs_sf_pl <- obs |>
  
  # Selecting the interesting fields
  
  transmute(
    id,
    taxon_name,                     
    observed_on = as.Date(observed_on),
    latitude,
    longitude,
    url_picture,
    uri,
    user_login
  ) |>
  
  # 2. Deleting rows without spatial coordinates
  
  filter(!is.na(latitude), !is.na(longitude)) |>
  
  # 3. Convert fiels to spatial points  WGS84. 
  
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326, remove = FALSE)

  # 4. Drawing map with leaflet

leaflet() |>
  addProviderTiles("Esri.WorldImagery", group = "Satélite") |>
  addProviderTiles("OpenStreetMap", group = "OSM") |>
  addCircleMarkers(
    data = obs_sf_pl,
    radius = 6,
    stroke = FALSE,          # sin borde
    fillOpacity = 0.9,
    fillColor = "red",
    popup = ~paste0(
      "<b>", taxon_name, "</b><br>",
      "Obs ID: ", id, "<br>",
      "User name: ", user_login, "<br>",
      "Date: ", observed_on, "<br>",
      "Specie: ", taxon_name, "<br>",
      "<img src='",url_picture, "' width='100' style='border-radius:5px; margin:1px 0;'>","<br>", # To show the picture
      "<a href='", uri, "' target='_blank'>View in Minka-sdg.com</a>"


    )
  ) |>
  addLayersControl(baseGroups = c("Satélite", "OSM"))

```

 - **<u>mnk_obs_byday</u>**
 
 ### - Auxiliary functions

 

