---
title: "Introduction to rminka"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to rminka}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup}
library(rminka)
```
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r carga, include=FALSE}
library(dplyr)
library(jsonlite)
library(httr)
library(leaflet)
#library(magick)
library(rminka)
library(tibble)
library(knitr)
```

## About

`rminka` is a wrapper for Minka APIs for accessing the observations.


## Installation

The easiest way to get `rminka` is to install from github repository with `pak`: 

```{r install, eval = FALSE, echo= TRUE}
pak::pak("Raiservi/rminka")
```

## Quickstart guide

These functions are designed to be used together. For queries that span multiple years, you can easily loop through the years of interest, run the appropriate function, and then combine the resulting tibbles with `dplyr::bind_rows()`.
> It is recommended to set the `quality` argument to `"research"` in order to 
get more reliable data that has been validated by several contributors.



A set of complementary  functions to find projects and their associated observations.

### - Project Queries 

- **<u>mnk_proj_byname</u>**

  Finds a project's ID using an approximate project name.
```{r project1, include = TRUE , echo=TRUE}
prj_names <- mnk_proj_byname("2025")

prj_names 

#Para hacer la tabla mas leible usamos `kable`

kable(prj_names[,c(1:2)])

```
- **<u>mnk_proj_info</u>**
 
   Retrieves detailed project information using its known ID.
```{r project_info, include = TRUE , echo=TRUE}

prj_info <- mnk_proj_info(420)

prj_info

```

It´s posible to known the users subscribed in this projects changing de parameter 'user = TRUE'  and the function gives a list of the id users.

```{r project_info_proj, include = TRUE , echo=TRUE}

prj_info_us <- mnk_proj_info(419, users = TRUE)

prj_info_us

```

  - **<u>mnk_proj_obs</u>**

Retrieves 

```{r project_obs:proj, include = TRUE , echo=TRUE}

prj_obs <- mnk_proj_obs(419, year= 2025, month=5)

prj_obs[2:14]

```


### - User Queries  
  
 - **<u>mnk_user_byname</u>**

```{r user_name, include = TRUE , echo=TRUE}

user_name <- mnk_user_byname("ramon")

user_name

```

 - **<u>mnk_user_proj</u>**
 
```{r user_project, include = TRUE , echo=TRUE}

user_project <- mnk_user_proj(6)

user_project

```

 - **<u> mnk_user_obs</u>**

```{r user_obs, include = TRUE , echo=TRUE}

user_obs <- mnk_user_obs(user_id=6, year = 2025, month = 8)

user_obs

```


### - Place Queries

 - **<u>mnk_places_byname</u>**

```{r places_byname, include = TRUE , echo=TRUE}

places <- mnk_places_byname("Forum")
places[,1:6]

```

 - **<u>mnk_places_sf</u>**

```{r places_sf, include = TRUE , echo=TRUE}

library(sf)
library(leaflet)


# 1. Downloading the geometry 
place <- mnk_place_sf(253)

# 2. Adding information from previos function
place <- place |>
  mutate(
    name = places[2,3],
    id = places[2,1]
  )

#Finding the map limits
b <- st_bbox(place)
xmin <- as.numeric(b[1]) # xmin
ymin <- as.numeric(b[2]) # ymin
xmax <- as.numeric(b[3]) # xmax
ymax <- as.numeric(b[4]) # ymax

# Drawing the map
leaflet(place) |>
  addProviderTiles("OpenStreetMap", group = "OSM") |>
  addProviderTiles("Esri.WorldImagery", group = "Satélite") |>
  addPolygons(
    color = "#2c4fb8",
    weight = 2,
    opacity = 1,
    fillOpacity = 0.4,
    label = ~name,
    popup = ~paste0("<b>", name, "</b><br>ID: ", id),
    highlightOptions = highlightOptions(weight = 3, bringToFront = TRUE)
  ) |>
  addLayersControl(baseGroups = c("Satélite", "OSM")) |>
  fitBounds(xmin, ymin, xmax, ymax)
```

 - **<u>mnk_places_obs</u>**
 
```{r places_obs, include = TRUE , echo=TRUE}

library(sf)
library(leaflet)

obs_place <- mnk_place_obs(place_id= 253, year = 2025, month = 2)

obs_place


obs_sf <- obs_place |>
  # 1. Selecting 
  transmute(
    id,
    taxon_name,                    # si no existe, créala antes con mutate(taxon_name = "...")
    observed_on = as.Date(observed_on),
    latitude,
    longitude,
    positional_accuracy
  ) |>
  # 2. elimina filas sin coordenadas
  filter(!is.na(latitude), !is.na(longitude)) |>
  # 3. convierte a puntos WGS84. remove = FALSE mantiene lat/lon como columnas
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326, remove = FALSE)

leaflet() |>
  addProviderTiles("Esri.WorldImagery", group = "Satélite") |>
  addProviderTiles("OpenStreetMap", group = "OSM") |>
  addPolygons(
    data = place,
    color = "#2c4fb8", weight = 2, fillOpacity = 0.4,
    label = ~name
  ) |>
  addMarkers(
    data = obs_sf,
    popup = ~paste0(
      "<b>", taxon_name, "</b><br>",
      "ID: ", id, "<br>",
      "Fecha: ", observed_on
    ),
    clusterOptions = markerClusterOptions()
  ) |>
  addLayersControl(baseGroups = c("Satélite", "OSM"))

```


### - Observation Queries

 - **<u>mnk_obs_id</u>**

```{r obs_id, include = TRUE , echo=TRUE}

obs_id <- mnk_obs_id(id = 553028)

obs_id

```

 - **<u>mnk_obs</u>**
 
```{r obs, include = TRUE , echo=TRUE}

#In this example don´t show messages in console (quiet= TRUE)

obs <- mnk_obs(taxon_name= "Diplodus sargus", year=2025, user_id=6,quiet = TRUE)

obs[,c(1,21,2,10,11,16,27)]


obs_sf_pl <- obs |>
  
  # Selecting the interesting fields
  
  transmute(
    id,
    taxon_name,                     
    observed_on = as.Date(observed_on),
    latitude,
    longitude,
    url_picture,
    uri,
    user_login
  ) |>
  
  # 2. Deleting rows without spatial coordinates
  
  filter(!is.na(latitude), !is.na(longitude)) |>
  
  # 3. Convert fiels to spatial points  WGS84. 
  
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326, remove = FALSE)

  # 4. Drawing map with leaflet

leaflet() |>
  addProviderTiles("Esri.WorldImagery", group = "Satélite") |>
  addProviderTiles("OpenStreetMap", group = "OSM") |>
  addCircleMarkers(
    data = obs_sf_pl,
    radius = 6,
    stroke = FALSE,          # sin borde
    fillOpacity = 0.9,
    fillColor = "red",
    popup = ~paste0(
      "<b>", taxon_name, "</b><br>",
      "Obs ID: ", id, "<br>",
      "User name: ", user_login, "<br>",
      "Date: ", observed_on, "<br>",
      "Specie: ", taxon_name, "<br>",
      "<img src='", url_picture, "' width='100' style='border-radius:5px; margin:1px 0;'>","<br>", # To show the picture
      "<a href='", uri, "' target='_blank'>View in Minka-sdg.com</a>"


    )
  ) |>
  addLayersControl(baseGroups = c("Satélite", "OSM"))

```

 - **<u>mnk_obs_byday</u>**
 
 ### - Auxiliary functions

 - **<u>get_wrm_tax</u>**
 
```{r get_tax, include = TRUE , echo=TRUE}



```
 
 - **<u>shrt_name</u>**
 

