---
title: "Introduction to rminka"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to rminka}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup}
library(rminka)
```
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r carga, include=FALSE}
library(dplyr)
library(jsonlite)
library(httr)
#library(leaflet)
#library(magick)
library(rminka)
library(tibble)
library(knitr)
```

## About

`rminka` is a wrapper for Minka APIs for accessing the observations.


## Installation

The easiest way to get `rminka` is to install from github repository with `pak`: 

```{r install, eval = FALSE, echo= TRUE}
pak::pak("Raiservi/rminka")
```

## Quickstart guide

These functions are designed to be used together. For queries that span multiple years, you can easily loop through the years of interest, run the appropriate function, and then combine the resulting tibbles with `dplyr::bind_rows()`.
> It is recommended to set the `quality` argument to `"research"` in order to 
get more reliable data that has been validated by several contributors.

### Project Queries

A set of complementary  functions to find projects and their associated observations.

**- Project Queries ** 

* mnk_proj_byname 
  Finds a project's ID using an approximate project name.
```{r project1, include = TRUE , echo=TRUE}
prj_names <- mnk_proj_byname("2025")

prj_names 

#In detail:

kable(prj_names[,c(1:5)])

```
* mnk_proj_info
   Retrieves detailed project information using its known ID.
```{r project_info, include = TRUE , echo=TRUE}

prj_info <- mnk_proj_info(420)

prj_info

kable(as.data.frame(prj_info))

```
It´s posible to known the users subscrived in this projects chnging de parameter 'user = TRUE'  and the function gives a list of the id users.

```{r project_info_users, include = TRUE , echo=TRUE}

prj_info_us <- mnk_proj_info(419, users = TRUE)

prj_info_us

kable(t(as.data.frame(prj_info_us)))

```

  
**- mnk_proj_obs **
   

kable(as.data.frame(obs_prj))
```

**- User Queries ** 
  
* mnk_user_byname 

```{r user_name, include = TRUE , echo=TRUE}

user_name <- mnk_user_byname("ramon")

user_name

kable(as.data.frame(user_name))
```

* mnk_user_proj 

```{r user_project, include = TRUE , echo=TRUE}

user_project <- mnk_user_proj(6)

user_project

kable(as.data.frame(user_project))
