---
title: "rminka"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include=FALSE}
library(dplyr)
library(jsonlite)
library(httr)
library(leaflet)
library(magick)
library(rminka)
```

<style>
.leaflet {
    margin: auto;
}
</style>



<style>
.picture {
    margin: auto;
}
</style>


## About

`rminka` is a wrapper for Minka APIs for accessing the observations.


## Installation

The easiest way to get `rminka` is to install from github repository: 

```{r install, eval = FALSE, echo= TRUE}
devtools::install_github("raiservi/rminka")
```

## Quickstart guide

### Get observations

`mnk_prj_obs()` This function retrieves observations from Minka. The text or taxon search can be refined by observation date, record quality and location.

> It is recommended to set the `quality` argument to `"research"` in order to get more reliable data that has been validated by several contributors.

#### Fuzzy search

You can search for observations by either common or scientific name. It will search the entire  database, so the search below will return entries that _mention_ seabream fishes , not just seabream observations.


```{r search0, include = TRUE , echo=TRUE}
sard <- mnk_obs_id(4251)

unique(sard$scientific_name)
```


> Note that `mnk_prj_obs()` will return 500 observations by default. This can be controlled with the `maxresults` argument.

```{r search1, include = TRUE , echo=TRUE}

sard_max_results <- mnk_proj_obs(420, year = 2025, month=5)

sard_max_results[,1:6]
```


